Use of significance test logic by scientists in a novel reasoning task

### 2020-11-04 inst/manuscript/manuscript.Rmd
Revision 1, RSOS (package)

abstract: |
  Although statistical significance testing is one of the most widely-used techniques across science, previous research has suggested that scientists have a poor understanding of how it works. If scientists misunderstand one of their primary inferential tools the implications are dramatic: potentially unchecked, unjustified conclusions and wasted resources. Scientists' apparent difficulties with significance testing have led to calls for its abandonment or increased reliance on alternative tools, which would represent a substantial, untested, shift in scientific practice. However, if scientists' understanding of significance testing is truly as poor as thought, one could argue such drastic action is required. We show using a novel experimental method that scientists do, in fact, understand the logic of significance testing and can use it effectively. This suggests that scientists may not be as statistically-challenged as often believed, and that reforms should take this into account.

@@

The interpretation of studies of researchers' understanding of SST are limited, however, by their methodology. A typical study presents a vignette describing research results. Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations that are thought to represent participants' understanding of SST. However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication). It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST or, say, familiarity with the technical terminology used to present statistical results. Conceptual understanding and fluency with common representations are both important, but are distinct.

A second major piece of evidence for misunderstandings of SST logic is reasoning errors in published papers [@Gelman:Stern:2006;@Hoekstra:etal:2006;@Nieuwenhuis:etal:2011;@Weisburd:etal:2003]. Like evidence from vignette studies, however, these errors are difficult to interpret as misunderstandings of SST logic *per se*. These examples show that whatever process lead to the statistical conclusion was flawed in some way, but many processes contribute to such conclusions. Cognitive (cites), technological (cites), and social processes (cites) have all been assigned some blame for statistical reasoning failures.

In deciding how to improve statistical reasoning, it is crucial to know where the problems are. The *fact* of reasoning problems tells us little about their *source*. In assessing potential interventions, however, the source is crucial. Some interventions might focus on the social aspects (e.g., decreasing the need for "significant" results for prestige), some on technological aspects (e.g., presenting statistical results in ways that were previously impossible), and some on cognitive aspects (e.g., adopting Bayesian procedures because these are claimed to be better understood).

To avoid conflating basic reasoning processes and fluency with common statistical terminology, we avoid using common statistics---or, indeed, any numbers---at all. Instead of focusing on familiar statistical language and tests participants' fluency with existing procedures, we adopt a different approach: we test working scientists' understanding of the basic conceptual framework underlying SST using a simulated experimental task.

The key innovation allowing us to focus on SST reasoning was to design an experiment that prevents the use of alternative strategies. A key feature of SST is that the use of a null distribution destroys information about effect sizes and sample size. In fact, this aspect of SST reasoning is often criticized. Alternative methodologies often focus on effect sizes (point estimates, confidence intervals, equivalence, likelihood, Bayesian priors/posteriors). We offered our participants only the information in a $p$ value, and participants had to understand or discover how to obtain that information. Their task was to use this information to come to a decision about the true sign of an effect through repeated experimentation.

If participants have poor understanding of SST, they would 1) often come to the wrong conclusion, in spite of ample information; 2) show error rates that are only weakly associated with true effect size; 3) be unable to articulate strategies for performing our task; 4) be sensitive to misleading, task-irrelevant information; 5) be insensitive to SST-relevant information. The scientists in our sample showed often came to the right conclusion, with their performance showing that they were sensitive to the information they were given. They explicitly reported using SST strategies. Our results suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.

@@

In tests of perception---e.g., of colorblindness [@Ishihara:1972]---it is common to eliminate one cue (e.g., brightness) in order assess the ability to use another (e.g., color). If color is the only useful cue for reading a number, deficits in color vision make the number difficult to read. We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution (i.e., SST logic). Without numerical information, many other strategies and heuristics, such as confidence intervals, or Bayesian inference, are difficult or impossible to apply.[^1]

@@

Participants were scientists or trainees recruited via social media. Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy. A demonstration version of the task that does not store data can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html. Because the study was run around the Christmas holiday season, it was hoped that the theme would make the task more enjoyable. The numerical information for an experiment, including sample size and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}). Participants could change the sample size per group for each experiment (increasing the time required to return a result), but did not know its numerical value. Importantly, the meaning of the colors and locations was unknown to the participants, aside from the monotone relationship with the sample size and statistical evidence.

@@

Participants were randomly assigned to one of 15 effect size conditions: either no difference ($\delta=0$), or $\delta=\pm0.1$, $\pm0.185$, $\pm0.296$, $\pm0.433$, $\pm0.596$,$\pm0.785$, or $\pm1$ standard deviation units. Each participant had a 25% probability of being assigned $\delta=0$, with the other 75% being randomly and uniformly distributed across the remaining 14 effect size conditions. These true effect sizes were not revealed to the participants. Their goal was to determine the sign of the effect (ie, which of the two teams is truly faster).

@@

@@

After sampling as many "experiments" and "random shuffle reports" as they liked, participants could report whether they believed Jinglies or Sparklies were better, that they could not detect a difference, that there was no difference, or that they were bored and wanted to stop. Following their decision they were asked several open-ended questions about their strategy. Our key questions are whether participants can effectively find the "truth", whether they report strategies consistent with SST, and whether their behaviour shows evidence of SST use.

@@

Here, we report the results of `r total_n` scientists or trainees who completed the statistical reasoning task. We provide an interactive app for exploring participants' responses at https://richarddmorey.shinyapps.io/explore/.

@@

Participants also made use of "replications" of the fictional experiments. Figure\@ref(fig:sampling)B shows the distribution of the number of experiments sampled as a function of the true effect size. Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`). When the effect size is small and difficult to detect, participants experimented more before deciding.

@@

When there was a true effect ($|\delta|>0$), correct decisions increased as a function of effect size, plateauing at about 95%. Of the `r total_n - null_n` participants for whom $|\delta|>0$, only `r sign_err_count` (`r round(sign_err_count/(total_n - null_n)*100,1)`%) indicated the incorrect team [a sign, or Type S, error; @Gelman:Carlin:2014]. For larger effect sizes, participants never incorrectly indicated that the two groups were the same.

Signal detection theory gives us another perspective on the error rates shown in Figure`~`{=latex}\ref{fig:error-rates}, allowing us to correct for the baseline of errors that occur in the null condition [@Macmillan:Creelman:2005]. We combine the "false alarm" rate when $\delta=0$ (`r round(t1_err_count/null_n * 100, 1)`%) with the "hit rates" for all other conditions using a simple signal detection model; see Supplement A, section 8 for model details. The fitted model yields $d'$ parameters that range from 1.41 when $\delta=0.1$ to 3.24 when $\delta=1$.

@@

Another way one can evaluate the participants' responses is whether they reflect the information in the display at the time the decision is made, taking into account all points. To roughly quantify the evidence for a difference for each participant, we computed two $p$ values from Wilcoxon tests using the fictitious experimental results as they stood when the participant made their decision: a signed-rank test on the experimental samples alone, and a rank-sum test between the shuffle reports and the experimental samples. These two $p$ values indicate the information available to participants using sign-like significance tests and those using the null samples, respectively. The rank-sum $p$ value was typically lower; although it makes little difference to the qualitative results, to fairly account for the information available to the participant, we used the smaller of the two $p$ values. In general, smaller $p$ values suggest a larger observed between the shuffle reports and the experiments, allowing us to compare the stimulus the participants were given to their decisions.

@@

Figure \ref{fig:response-ev} shows the distribution of Wilcoxon $p$ values (arranged by the direction of the decision). Kernel density estimates show the distributions of $p$ values when participants indicated that Sparklies were faster, no detection/same, or that Jinglies were faster. With a few notable exceptions, participants' conclusions appear reasonable given the information in the display, though a few participants appear to ignore clear evidence of an effect.

@@

In addition to a random effect size, participants were randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic. Of particular interest was how the transformation affected responding for the same visual deviation from the center.

@@

If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider. When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant, given the $p$ value.

@@

Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$). It is noteworthy that when they responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with some participants incorrectly using the non-diagnostic visual extremeness to perform the task. If more people had been fooled by the irrelevant visual manipulation of the sampling distribution, we would expect this effect to be substantially larger.

@@

We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategies), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports.

@@

As Table \ref{tab:strategies} shows that a large majority of participants (`r for_table[1,1]`, `r round( 100*as.numeric(for_table[1,1]) / as.numeric(for_table[1,6]), 2)`%) indicated using strong significance testing strategies. We should be cautious in directly interpreting this high number alone, however, because participants were told in the instructions that the shuffle reports could be used for assessing sampling variability. We did this to make clear what the shuffle reports were, but without explaining how to use them. To some extent, then, the text responses may reflect the instructions. However, the data strongly suggest a deeper understanding; first, among the responses were richer, lucid descriptions of SST logic, such as:

@@

Secondly---and most importantly---the instructions did not tell the participants *how* they should use the shuffles reports, yet many participants gave detailed accounts. Combined with the other reported results, this strongly suggests that our participants---with some exceptions---do understand the basic SST logic and can deploy it to correctly solve novel problems.

@@

Although it has previously been suggested that scientists have dramatic misunderstandings of SST logic, scientists and trainees in our experiment demonstrate both understanding and the ability to use it to come to the correct conclusion in a simulated statistical task. Moreover, they report strategies consistent with SST, and the signature of SST reasoning can be seen in their responses. Because we removed numerical effect size and sample size information --- making strategies other than pure significance testing difficult or impossible to apply --- our results are evidence that scientists *can* successfully deploy SST logic. It is still an open question what causes SST to be misunderstood so often, but it seems comprehension of its underlying logic can be ruled out for many scientists.

Our findings echo other work showing that human reasoning can, under some conditions, be better than previously understood. [@Cosmides:Tooby:1992;@Gigerenzer:Hoffrage:1995]. Suggestions that SST be discontinued due to scientists' apparent misunderstandings may have been hasty. Of course, there may be other reasons to abandon SST, but our work shows that given the opportunity, scientists successfully deploy basic SST logic. In spite of scientists' real-life statistical behaviour often resembling a "ritual" [@Gigerenzer:etal:2004], when we eliminate the ritual --- no $p$ value, or any other familiar number, was offered --- they think statistically, very often arriving at the correct conclusion about the sign of the effect.

We wish to emphasize what we cannot, and do not, argue. First, we cannot argue that simply because scientists can successfully use SST logic, that they do in real situations, or that specific instantiations of SST, such as $p$ values, are used well. We specifically set out to abstract the logic away from the typical ways and situations in which scientists use the logic. This has the benefit of being helping to identify where problems might be, but the downside that generalizing the results will require further work. We are also not addressing other potential arguments against SST, such as philosophical ones.

Although our conclusions are different those of other studies, there are some major differences between our methods and the methods of previous demonstrations that scientists fail when reasoning about SST. We have already mentioned the fact that previous demonstrations typically use survey vignettes, as opposed to our more engaged experimental method. A second difference is our sample which is a substantially larger and more diverse convenience sample than previous demonstrations. It is possible that the recent so-called "replication crisis" [@Baker:2016] has raised awareness among scientists of deficits of statistical reasoning. If this increase in awareness accounts for our more optimistic findings, this would be good news for statistical educators.

Finally, we hope to provide a fresh method and perspective on a long-standing debate in statistical cognition. Simulation-based approaches to teaching statistics have long been touted [@Cumming:etal:1995;@Rossman:Chance:2014]. Simulation-based approaches to *studying* scientists' statistical reasoning may also profitable, particularly in studying reasoning that is difficult for participants to articulate formally. If we are to reform statistical education and practice in the sciences, we should base that reform on diverse lines of evidence about scientists' reasoning. Harnessing scientists' already-existing statistical competence may make effective methodological reform in the sciences easier than wholesale abandoning of SST.

>>>>

abstract: |
  Although statistical significance testing is one of the most widely-used techniques across science, previous research has suggested that scientists have a poor understanding of how it works. If scientists misunderstand one of their primary inferential tools the implications are dramatic: potentially unchecked, unjustified conclusions and wasted resources. Scientists' apparent difficulties with significance testing have led to calls for its abandonment or increased reliance on alternative tools, which would represent a substantial, untested, shift in scientific practice. However, if scientists' understanding of significance testing is truly as poor as thought, one could argue such drastic action is required. We present evidence using a novel reasoning task that scientists may understand the logic of significance testing better than previously thought. Scientists may not be as statistically-challenged as often believed; reforms should take this into account.

@@

The interpretation of studies of researchers' understanding of SST is limited, however, by their methodology. A typical study presents a vignette describing research results. Statistical results are offered to the participants (e.g., a $t$ statistic and $p$ value), who are then asked to explicitly give or endorse various interpretations. These responses are taken to represent participants' understanding, or misunderstanding, of SST. However, there are reasons to be cautious of drawing strong conclusions from these studies, including the abstract nature of such vignettes, the lack of investment researchers have in the fictional research, and their disconnection from research activity (e.g., experimentation and replication). It is unclear how well vignette studies (including ones by the present authors: [@Hoekstra:etal:2014;@Hoekstra:etal:2012]) tap understanding of the core logic of SST rather than, say, familiarity with the technical terminology used to present statistical results. Conceptual understanding and fluency with common representations are both important, but are distinct.

A second major piece of evidence for misunderstandings of SST logic is reasoning errors in published papers [@Gelman:Stern:2006;@Hoekstra:etal:2006;@Nieuwenhuis:etal:2011;@Weisburd:etal:2003]. Like evidence from vignette studies, however, these errors are difficult to interpret as misunderstandings of SST logic *per se*. These examples show that whatever process lead to the statistical conclusion was flawed in some way, but many processes contribute to such conclusions. Cognitive [e.g., @Pashler:Harris:2012], technological [e.g., @Kennedy-Shaffer:2019], and social processes [e.g., @Lilienfeld:2017] have all been assigned some blame for statistical reasoning failures.

In deciding how to improve statistical reasoning, it is crucial to know where the problems lie. The *fact* of reasoning problems tells us little about their *source*. In assessing potential interventions, however, the source is crucial. Some interventions might focus on the social aspects (e.g., decreasing the need for "significant" results for prestige), some on technological aspects (e.g., presenting statistical results in ways that were previously impossible), and some on cognitive aspects (e.g., adopting Bayesian procedures because these are claimed to be better understood).

To avoid conflating basic reasoning failures and lack of fluency with common statistical terminology, we avoid using common statistics---or, indeed, any numbers---at all. Instead of focusing on familiar statistical language and tests participants' fluency with existing procedures, we adopt a different approach: we test working scientists' understanding of the basic conceptual framework underlying SST using a simulated experimental task.

The key innovation allowing us to focus on SST reasoning was to design an experiment that prevents the use of alternative strategies. A critical feature of SST is that the use of a null distribution destroys information about effect sizes and sample size. In fact, this aspect of SST reasoning is often criticized, while alternative methodologies focus on effect sizes (point estimates, confidence intervals, equivalence, likelihood, Bayesian priors/posteriors). We offered our participants only the information in a $p$ value, and participants had to understand or discover how to obtain that information. Their task was to use this information to come to a decision about the true sign of an effect through repeated experimentation.

If participants have poor understanding of SST, they would 1) often come to the wrong conclusion, in spite of ample information; 2) show error rates that are only weakly associated with true effect size; 3) be unable to articulate strategies for performing our task; 4) be sensitive to misleading, task-irrelevant information; 5) be insensitive to SST-relevant information. The scientists in our sample often came to the right conclusion, and their performance showing sensitivity to the SST-relevant information they were given. Moreover, they explicitly reported using SST strategies. Our results suggest that common methods for assessing scientists' competence may miss important aspects of their statistical knowledge, and hence that the case for abandoning significance testing may be overstated.

@@

In tests of perception, it is common to eliminate one cue in order assess the ability to use another: e.g., eliminating brightness cues to test colorblindness [@Ishihara:1972]. If color is the only useful cue for reading a number on a card, deficits in color vision make the number difficult to read. We adopt a similar strategy to test statistical reasoning: we eliminate numerical information from statistical results to test scientists' ability to interpret results with reference to a null sampling distribution, a central element of SST logic. Without numerical information, many other strategies and heuristics, such as confidence intervals, or Bayesian inference, are difficult or impossible to apply.[^1]

@@

Participants were scientists or trainees recruited via social media. Our statistical reasoning task required them to perform a series of experiments to judge which of two groups of "Christmas elves" --- "Jinglies" or "Sparklies" --- could make more of a particular toy. A demonstration version of the task can be found at https://richarddmorey.github.io/Morey_Hoekstra_StatCognition/articles/task_demo.html. Because the study was run around the Christmas holiday season, we hoped the theme would make the task more engaging. The numerical information for an experiment, including sample size and the test statistic, was translated into color and location and displayed as a point on a two-dimensional visual interface (Figure \ref{fig:interface}). Participants could change the sample size per group for each experiment (increasing the time required to return a result), but did not know its numerical value. Importantly, the meaning of the colors and locations was unknown to the participants, aside from the monotone relationship with the sample size and statistical evidence.

@@

Participants were randomly assigned to one of 15 effect size conditions: either no difference ($\delta=0$), or $\delta=\pm0.1$, $\pm0.185$, $\pm0.296$, $\pm0.433$, $\pm0.596$,$\pm0.785$, or $\pm1$ standard deviation units. Each participant had a 25% probability of being assigned $\delta=0$, with the other 75% being randomly and uniformly distributed across the remaining 14 effect size conditions. These true effect sizes were not revealed to the participants. Their goal was to determine the sign of the effect (i.e., which of the two teams is truly faster).

@@

```{r}
MoreyHoekstra2019::christmas_stats_participants %>%
  group_by(shuffle) %>%
  summarise(n=n()) %>%
  mutate(p = n / sum(n),
         yes = grepl(pattern = "yes", shuffle)) -> shuffle_understanding

shuffle_understanding %>% filter(yes) %>%
  pull(n) -> shuffle_ns

shuffle_understanding %>% filter(yes) %>%
  pull(p) -> shuffle_ps

shuffle_pct = round(shuffle_ps * 100,1)

@@

After sampling as many "experiments" and "random shuffle reports" as they liked, participants could report whether they believed Jinglies or Sparklies were the better team, that they could not detect a difference, that there was no difference, or that they were bored and wanted to stop. Following their decision they were asked several open-ended questions about their strategy, along with some opinion and demographic questions. Our central questions are whether participants can effectively find the "truth", whether they report strategies consistent with SST, and whether their behaviour shows evidence of strategic SST use.

@@

Here, we report the results of `r total_n` scientists or trainees who completed the statistical reasoning task.

@@

Participants also made use of "replications" of the fictional experiments. Figure`~`{=latex}\@ref(fig:sampling)B shows the distribution of the number of experiments sampled as a function of the true effect size. Median numbers of experiments range from 20 when Jinglies and Sparklies were equally fast, down to 9 when the true effect size was $\delta=1$ and thus the effect was relatively easy to detect (`r papaja::apa_print(kt_expt)$statistic`). When the effect size is small and difficult to detect, participants experimented more before deciding.

@@

When there was a true effect ($\delta\neq0$), correct decisions increased as a function of effect size, plateauing at about 95%. Of the `r total_n - null_n` participants for whom $\delta\neq0$, only `r sign_err_count` (`r round(sign_err_count/(total_n - null_n)*100,1)`%) indicated the incorrect team [a sign, or Type S, error; @Gelman:Carlin:2014]. For larger effect sizes, participants never incorrectly indicated that the two groups were the same.

Signal detection theory gives us another perspective on the decision rate shown in Figure`~`{=latex}\ref{fig:error-rates}, allowing us to correct for the baseline of errors that occur in the null condition [@Macmillan:Creelman:2005]. We combine the "false alarm" rate when $\delta=0$ (`r round(t1_err_count/null_n * 100, 1)`%) with the "hit rates" for all other conditions using a simple signal detection model; see Supplement A, section 8 for model details. The fitted model yields $d'$ parameters that range from 1.41 when $\delta=0.1$ to 3.24 when $\delta=1$.

@@

Another way of evaluating participants' responses is whether they reflect the information in the display at the time the decision is made, taking into account all points. To roughly quantify the evidence for a difference for each participant, we computed two $p$ values from Wilcoxon tests using the fictitious experimental results as they stood when the participant made their decision: a signed-rank test on the experimental samples alone, and a rank-sum test between the shuffle reports and the experimental samples. These two $p$ values indicate the information available to participants using sign-like significance tests and those using the null samples, respectively. The rank-sum $p$ value is based on more information and so was typically lower. It makes little difference to the qualitative results, but to fairly account for the information available to the participant, we used the smaller of the two $p$ values. In general, smaller $p$ values suggest a larger observed between the shuffle reports and the experiments, allowing us to compare the stimulus the participants were given to their decisions.

@@

Figure \ref{fig:response-ev} shows the distribution of Wilcoxon $p$ values (arranged by the direction of the decision). Kernel density estimates show the distributions of $p$ values when participants indicated that Sparklies were faster, no detection/same, or that Jinglies were faster. With a few notable exceptions, participants' conclusions appear reasonable given the information in the display, though a few participants appear to ignore clear evidence of an effect. We provide an interactive app for exploring participants' individual responses at https://richarddmorey.shinyapps.io/explore/.

@@

In addition to a random effect size, participants were also randomly assigned to one of two transformations of the location/color test statistic from an underlying $z$ statistic. Of particular interest was how the transformation affected responding for the same visual deviation from the center.

@@

If participants were using the shuffle reports to interpret the data, as would be predicted if they were using SST logic, the transformation should affect their interpretation of the visual evidence: a visually-extreme point should be more discounted against the sampling distribution that is wider. When we break down responses by the *visual* extremeness of the evidence, responses in two conditions should appear different; when we break down responses by  *statistical* extremeness (i.e., the $p$ value) responses in the two conditions should appear very similar, because the visual manipulation is irrelevant given the $p$ value.

@@

Applying the same analysis to the responses corrected for their respective sampling distributions (Figure \ref{fig:response-ev2}, bottom) almost completely eliminates the effect of experimental condition, as would be expected if most participants were using the sampling distributions to calibrate (sequential LRT: $\chi^2_2 = `r round(sum(glm_anova_p[3:4,3]),3)`, p=`r round(1 - pchisq(sum(glm_anova_p[3:4,3]),2),3)`$). It is noteworthy that when the responses are aligned by sampling distribution, the wide condition appears to slightly dominate; this is consistent with some participants incorrectly using the non-diagnostic visual extremeness to perform the task. If more people had been fooled by the irrelevant width of the null sampling distribution, we would expect this effect to be substantially larger.

@@

We coded their responses according to whether they indicated comparing to the shuffle reports or using them to assess sampling variability (which we term "strong" significance testing strategies), assessing asymmetry in the display (a "weak" significance testing strategy, because it ignores information), and whether they explicitly deny using the shuffle reports (see Supplement`~`{=latex}A, section 7 for coding details).

@@

As Table \ref{tab:strategies} shows, a large majority of participants (`r for_table[1,1]`, `r round( 100*as.numeric(for_table[1,1]) / as.numeric(for_table[1,6]), 2)`%) indicated using strong significance testing strategies. We should be cautious in directly interpreting this high number alone, however, because participants were told in the instructions that the shuffle reports could be used for assessing sampling variability. We did this to make clear what the shuffle reports were, but without explaining *how* to use them. To some extent, then, the text responses may reflect the instructions. However, the data strongly suggest a deeper understanding; first, among the responses were richer, lucid descriptions of SST logic, such as:

@@

Secondly---and most importantly---the instructions did not tell the participants how they should use the shuffles reports, yet many participants gave detailed accounts. Combined with the other reported results, this strongly suggests that our participants---with some exceptions---do understand the basic SST logic and can deploy it to correctly solve novel problems.

@@

Although it has previously been suggested that scientists have dramatic misunderstandings of SST logic, scientists and trainees in our experiment demonstrate both understanding and the ability to use the logic to come to the correct conclusion in a simulated statistical task. Moreover, they report strategies consistent with SST, and signatures of SST reasoning can be seen in their responses. Because we removed numerical effect size and sample size information --- making strategies other than pure significance testing difficult or impossible to apply --- our results are evidence that scientists *can* successfully deploy SST logic. It is still an open question what causes typical SST reports to be misunderstood so often, but we have not found evidence that the problem is misapprehension of its underlying logic.

Our findings echo other demonstrations that human reasoning can, under some conditions, be better than previously understood. [@Cosmides:Tooby:1992;@Gigerenzer:Hoffrage:1995]. Suggestions that SST be discontinued due to scientists' apparent misunderstandings may be hasty. Of course, there may be other reasons to abandon SST, but our work shows that given the opportunity, scientists successfully deploy basic SST logic. In spite of scientists' real-life statistical behaviour often resembling a "ritual" [@Gigerenzer:etal:2004], when we eliminate the ritual --- no $p$ value, or any other familiar number, was offered --- they think statistically, very often arriving at the correct conclusion about the sign of the effect.

We wish to emphasize what we cannot, and do not, argue. First, we cannot argue that simply because scientists can successfully use SST logic, that they *do* in real situations, or that specific instantiations of SST, such as $p$ values, are used well. We specifically set out to abstract the logic away from the typical situations in which scientists use the logic. This has the benefit of being helping to identify where problems might be, but the downside that generalizing the results will require further work. We also cannot address other potential arguments against SST, such as philosophical ones.

Finally, we hope to provide a fresh method and perspective on a long-standing debate in statistical cognition. Simulation-based approaches to teaching statistics have long been touted [@Cumming:etal:1995;@Rossman:Chance:2014]. Simulation-based approaches to *studying* scientists' statistical reasoning may also profitable, particularly in studying reasoning that is difficult for participants to articulate formally. If we are to reform statistical education and practice in the sciences, we should base that reform on diverse lines of evidence about scientists' reasoning. Understanding and harnessing scientists' already-existing competence in statistical reasoning is essential to developing effective methodological reforms.
